### 2021-01-26 Phys 311/main.tex
Upload

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Academic Title Page
% LaTeX Template
% Version 2.0 (17/7/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% WikiBooks (LaTeX - Title Creation) with modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
% 
% Instructions for using this template:
% This title page is capable of being compiled as is. This is not useful for 
% including it in another document. To do this, you have two options: 
%
% 1) Copy/paste everything between \begin{document} and \end{document} 
% starting at \begin{titlepage} and paste this into another LaTeX file where you 
% want your title page.
% OR
% 2) Remove everything outside the \begin{titlepage} and \end{titlepage}, rename
% this file and move it to the same directory as the LaTeX file you wish to add it to. 
% Then add \input{./<new filename>.tex} to your LaTeX file where you want your
% title page.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters

\usepackage{mathpazo} % Palatino font
\usepackage{graphicx}
\usepackage{varioref}
\usepackage{float}

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage} % Suppresses displaying the page number on the title page and the subsequent page counts as page 1
	\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for horizontal lines, change thickness here

	\center % Centre everything on the page

	%------------------------------------------------
	%	Headings
	%------------------------------------------------

	\textsc{\LARGE University of Washington, Bothell}\\[1.5cm] % Main heading such as the name of your university/college

	\textsc{\Large CTA Term Paper}\\[0.5cm] % Major heading such as project name

	\textsc{\large B Phys 311, Intro to Astrophysics}\\[0.5cm] % Minor heading such as course title

	%------------------------------------------------
	%	Title
	%------------------------------------------------

	\HRule\\[0.4cm]

	{\huge\bfseries Segue 1: An Unevolved Fossil Galaxy from the Early Universe}\\[0.4cm] % Title of your document

	\HRule\\[1.5cm]

	%------------------------------------------------
	%	Author(s)
	%------------------------------------------------

	\begin{minipage}{0.4\textwidth}
		\begin{flushleft}
			\large
			\textit{Pongpak Techagumthorn}\\
			\textsc{pongpakt@uw.edu} % Your name
		\end{flushleft}
	\end{minipage}

	%------------------------------------------------
	%	Date
	%------------------------------------------------

	\vfill\vfill\vfill % Position the date 3/4 down the remaining page

	{\large January 25, 2021} % Date, change the \today to a set date if you want to be precise

	%------------------------------------------------
	%	Logo
	%------------------------------------------------

	%\vfill\vfill
	%\includegraphics[width=0.2\textwidth]{placeholder.jpg}\\[1cm] % Include a department/university logo - this will require the graphicx package

	%----------------------------------------------------------------------------------------

	\vfill % Push the date up 1/4 of the remaining page

\end{titlepage}

%----------------------------------------------------------------------------------------

\section{Description}
This is a test
Hiring Professionals (KajLagbe) is a platform for publishing and finding jobs that are not so conventional to the existing job marketplace. We focus on providing manpower specially for the day to day requirements.
\newline 
\newline
So, there are specially two types of user that will use this platform. One of them will be the Professionals with actual skills and the other type of the user will be the Consumer who will consume the service.
\newline
\newline
Consumers can post jobs specifying the skill that is needed for the job.
Professionals can provide their skills during registration, which will be the key to finding all of the available jobs. Professionals can also search jobs based on location.
\newline
\newline
When professionls see a job post s/he can comment on the post to seek the attention of the consumer. Then comment is posted successfully s/he will be regarded as an interested user and option will be available to consumer to submit contract to the professional. Multiple professionals can be interested. Consumer can choose among them based on their rating / skill / daily charge.
\newline
\newline
When contract is submitted to a professional, the contract will be visible to the contracts page of the professional. If a professional accepts the contract the job will be restricted for all other users to accept.
\newline
\newline
When a professional completes a job he will have to press Job Completed button which will generate the receipt with the bill. Professionals can then rate the consumer or return to home page. In the same way consumers will also get an option to rate the professional who has completed his job.

\clearpage

\section{Motivation}
The motivation behind this project was the thought that, if there was a platform where we can find the workers needed for our day to day living. If we needed an electrician we don't have to go looking for an electrician but the only thing that we have to do is just make a post, where many professionals (electrician) can find you and have your work done with the best price.
\clearpage

\section{Novelty}

What is new about this project?

\begin{itemize}
\item Get jobs that are posted daily
\item Get only the job suggestions that match your skills
\item Manage post and contracts more efficiently
\item Bill is calculated on daily basis
\end{itemize}

\clearpage

\section{Sample Picture}
\includegraphics[scale=0.5]{Untitled+design+(2).png}

\clearpage

\section{Tools Used}

\begin{table}[H]
\centering
\caption{Tools used for the project}
\label{my-label}
\begin{tabular}{|l|l|}
\hline
\textbf{Tool} & \textbf{Purpose}         \\ \hline
Xampp         & For creating a localhost \\ \hline
phpMyAdmin    & For hosting the database \\ \hline
MySQL    	  & As a database system \\ \hline
PHP    		  & For server side handling \\ \hline
Javascript    & For AJAX call and form validation \\ \hline
Bootstrap     & For responsive interface \\ \hline
CSS \& HTML   & For designing basic template \\ \hline
\end{tabular}
\end{table}

\clearpage

\section{Challenges Faced}
These are the challenges we faced while doing the project:
\begin{itemize}
\item \textbf{Challenge 1: } While implementing auto-complete we couldn't get data dynamically.
\item \textbf{Challenge 2: } Mailing with PHPMailer didn't work at first due to authorization problem and some protocols.
\item \textbf{Challenge 3: } Sending/receiving data in the background with AJAX was quite confusing at first.
\item \textbf{Challenge 4: } Calculating fare based on time interval was tough to implement at first.
\end{itemize}

\clearpage

\section{Applications}

Applications of our project is as follows:
\begin{itemize}
\item \textbf{Application 1: } Find jobs based on skills and place.
\item \textbf{Application 2: } Get suggestions of daily jobs that match your skills.
\item \textbf{Application 3: } Get the list of all of your contracts that are currently in place.
\item \textbf{Application 4: } Keep track of your completed jobs along with the earnings.
\item \textbf{Application 5: } Posting jobs that requires a specific skill
\item \textbf{Application 6: } Get the list of top employers
\item \textbf{Application 7: } Generate receipt based on the time duration of a service
\item \textbf{Application 8: } Comment on posts published by consumers
\item \textbf{Application 9: } Rating enabled for consumers and professionals when job is finished
\item \textbf{Application 10: } Each user have their own profile with all of their basic information, professionals have their skills also

\end{itemize}

\clearpage

\section{Future Work}
\begin{itemize}
\item Develop a mobile application based on this project.
\item Adding support for map for direction and estimated time of arrival.
\item One click to call feature.
\item Enabling professionals active time period.
\item Enabling individual messaging and replying to comments.
\item Adding automated fare calculation based on some factor (e.g. time, distance).
\item Discount will be provided to consumers.
\end{itemize}

\clearpage

\end{document}