### 2021-01-30 Phys 311/CTA Paper/main.tex
Move folder

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Academic Title Page
% LaTeX Template
% Version 2.0 (17/7/17)
%
% This template was downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% WikiBooks (LaTeX - Title Creation) with modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
% 
% Instructions for using this template:
% This title page is capable of being compiled as is. This is not useful for 
% including it in another document. To do this, you have two options: 
%
% 1) Copy/paste everything between \begin{document} and \end{document} 
% starting at \begin{titlepage} and paste this into another LaTeX file where you 
% want your title page.
% OR
% 2) Remove everything outside the \begin{titlepage} and \end{titlepage}, rename
% this file and move it to the same directory as the LaTeX file you wish to add it to. 
% Then add \input{./<new filename>.tex} to your LaTeX file where you want your
% title page.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[11pt]{article}

\usepackage[utf8]{inputenc} % Required for inputting international characters
\usepackage[T1]{fontenc} % Output font encoding for international characters

\usepackage{mathpazo} % Palatino font
\usepackage{graphicx}
\usepackage{varioref}
\usepackage{float}

\begin{document}

%----------------------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------------------

\begin{titlepage} % Suppresses displaying the page number on the title page and the subsequent page counts as page 1
	\newcommand{\HRule}{\rule{\linewidth}{0.5mm}} % Defines a new command for horizontal lines, change thickness here

	\center % Centre everything on the page

	%------------------------------------------------
	%	Headings
	%------------------------------------------------

	\textsc{\LARGE University of Washington, Bothell}\\[1.5cm] % Main heading such as the name of your university/college

	\textsc{\Large CTA Proposal}\\[0.5cm] % Major heading such as project name

	\textsc{\large B Phys 311, Intro to Astrophysics}\\[0.5cm] % Minor heading such as course title

	%------------------------------------------------
	%	Title
	%------------------------------------------------

	\HRule\\[0.4cm]

	{\huge\bfseries Segue 1: An Unevolved Fossil Galaxy from the Early Universe}\\[0.4cm] % Title of your document

	\HRule\\[1.5cm]

	%------------------------------------------------
	%	Author(s)
	%------------------------------------------------

	\begin{minipage}{0.4\textwidth}
		\begin{flushleft}
			\large
			\textit{Pongpak Techagumthorn}\\
			\textsc{pongpakt@uw.edu} % Your name
		\end{flushleft}
	\end{minipage}

	%------------------------------------------------
	%	Date
	%------------------------------------------------

	\vfill\vfill\vfill % Position the date 3/4 down the remaining page

	{\large January 25, 2021} % Date, change the \today to a set date if you want to be precise

	\vfill % Push the date up 1/4 of the remaining page

\end{titlepage}

%----------------------------------------------------------------------------------------

\section{Topic Background}

\section{Motivation}
The motivation behind this project was the thought that, if there was a platform where we can find the workers needed for our day to day living. If we needed an electrician we don't have to go looking for an electrician but the only thing that we have to do is just make a post, where many professionals (electrician) can find you and have your work done with the best price.
\clearpage

\section{Novelty}

What is new about this project?

\begin{itemize}
\item Get jobs that are posted daily
\item Get only the job suggestions that match your skills
\item Manage post and contracts more efficiently
\item Bill is calculated on daily basis
\end{itemize}

\clearpage

\section{Sample Picture}
\includegraphics[scale=0.5]{Untitled+design+(2).png}

\clearpage

\section{Tools Used}

\begin{table}[H]
\centering
\caption{Tools used for the project}
\label{my-label}
\begin{tabular}{|l|l|}
\hline
\textbf{Tool} & \textbf{Purpose}         \\ \hline
Xampp         & For creating a localhost \\ \hline
phpMyAdmin    & For hosting the database \\ \hline
MySQL    	  & As a database system \\ \hline
PHP    		  & For server side handling \\ \hline
Javascript    & For AJAX call and form validation \\ \hline
Bootstrap     & For responsive interface \\ \hline
CSS \& HTML   & For designing basic template \\ \hline
\end{tabular}
\end{table}

\clearpage

\section{Challenges Faced}
These are the challenges we faced while doing the project:
\begin{itemize}
\item \textbf{Challenge 1: } While implementing auto-complete we couldn't get data dynamically.
\item \textbf{Challenge 2: } Mailing with PHPMailer didn't work at first due to authorization problem and some protocols.
\item \textbf{Challenge 3: } Sending/receiving data in the background with AJAX was quite confusing at first.
\item \textbf{Challenge 4: } Calculating fare based on time interval was tough to implement at first.
\end{itemize}

\clearpage

\section{Applications}

\clearpage

\section{Future Work}
\begin{itemize}
\item Develop a mobile application based on this project.
\item Adding support for map for direction and estimated time of arrival.
\item One click to call feature.
\item Enabling professionals active time period.
\item Enabling individual messaging and replying to comments.
\item Adding automated fare calculation based on some factor (e.g. time, distance).
\item Discount will be provided to consumers.
\end{itemize}

\clearpage

\end{document}